---
title: "Mopeia surveillance dashboard"
author: joe@databrew.cc
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
---
  

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
library(knitr)
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r, results = 'hide'}
# Packages
library(ggplot2)
library(cism)
library(rgeos)
library(maptools)
library(rgdal)
library(tidyr)
library(RColorBrewer)
library(dplyr)
library(leaflet)
library(readr)
library(ggthemes)
library(ggrepel)
library(knitr)
library(kableExtra)
options(knitr.table.format = "html") 
# Define a theme
theme_mopeia <- ggthemes::theme_hc
library(flexdashboard)

```

```{r}
# Get data
source('get_data.R')
```

Overview
=====================================  


Row
-----------------------------------------------------------------------

### ACD visits carried out

```{r}
valueBox(length(which(visit_dates$visit_done)), icon = "fa-stethoscope")
```

### Malaria cases detected through ACD

```{r}
valueBox(length(which(visit_dates$malaria)), icon = "fa-flag",
         color = 'info')
```

### Children-years observed in ACD

```{r}
years <- round(nrow(etar) / 365.25)
valueBox(round(nrow(etar) / 365.25), 
         icon = "fa-binoculars",
         color = 'warning')
```


Row
-----------------------------------------------------------------------

### PCD visits among 0-4 year olds

```{r}
valueBox(nrow(pcd), icon = "fa-vcard", color = 'warning')
```

### PCD malaria cases among 0-4 year olds

```{r}
valueBox(length(which(pcd$malaria)), icon = "fa-vcard", color = 'danger')
```

### RDTs administered in PCD to 0-4 year olds

```{r}
valueBox(length(which(!is.na(pcd$rdt))), icon = "fa-vcard", color = 'info')
```


    
Row
-------------------------------------
    
### Study area

```{r}
leaflet() %>%
  # addTiles() %>%
  addProviderTiles('Esri.WorldImagery') %>%
  addPolygons(data = cism::mop2, color = 'darkorange')
```

### Village centroids

```{r}
village <- master_table %>%
  group_by(village_number) %>%
  summarise(children = n(),
            lat = mean(lat),
            lng = mean(lng),
            status = ifelse(dplyr::first(status), 'Spray', 'No spray')) %>%
  mutate(color = ifelse(status == 'Spray', 'darkorange', 'blue'))

leaflet() %>%
  # addTiles() %>%
  addProviderTiles('Stamen.TonerHybrid') %>%
  # addPolygons(data = cism::mop2) %>%
  addCircles(data = village,
                   lng = ~lng,
                   lat = ~lat,
                   color = ~color,
             popup = paste0('Village ', village$village_number, ': ', village$status))
```

ACD
=====================================  

    
    
Row
-------------------------------------
    
### Round-specific incidence figure

```{r}
# ACD ########################
# Round Incidence (seven of them)
# (# of RDT+ that round* / # of children in cohort that round)
#   * Round: equals visits of ACD. We are currently at the end of the 8th monthly follow up
x <- visit_dates %>%
  group_by(round) %>%
  summarise(children = n(),
            positives = length(which(malaria)),
            negatives = length(which(!malaria)),
            tested = length(!is.na(malaria))) %>%
  ungroup %>%
  mutate(percent_positive = positives / tested * 100) %>%
  mutate(Tests = tested)

ggplot(data = x,
       aes(x = round,
           y = percent_positive)) +
  geom_line(alpha = 0.7) +
  geom_point(alpha = 0.6,
             aes(size = Tests)) +
  labs(x = 'Round',
       y = 'Percent positive') +
  theme_mopeia() +
  scale_x_continuous(breaks = 1:max(x$round))
```

### Round-specific incidence table

```{r}
x <- x %>%
  dplyr::select(round, Tests, positives, negatives, percent_positive)
prettify(x)
```

Row
-------------------------------------
   
### Round incidence by spray status figure

```{r}
# ACD ########################
# Round Incidence (seven of them)
# (# of RDT+ that round* / # of children in cohort that round)
#   * Round: equals visits of ACD. We are currently at the end of the 8th monthly follow up
x <- visit_dates %>%
  group_by(round, spray_status) %>%
  summarise(children = n(),
            positives = length(which(malaria)),
            negatives = length(which(!malaria)),
            tested = length(!is.na(malaria))) %>%
  ungroup %>%
  mutate(percent_positive = positives / tested * 100) %>%
  mutate(Tests = tested)

ggplot(data = x,
       aes(x = round,
           y = percent_positive,
           color = spray_status)) +
  geom_line(alpha = 0.7) +
  geom_point(alpha = 0.6,
             aes(size = Tests)) +
  labs(x = 'Round',
       y = 'Percent positive') +
  theme_mopeia() +
  scale_x_continuous(breaks = 1:max(x$round)) +
  scale_color_manual(name= '',
                     values = c('darkorange', 'blue'))
```

### Round incidence by spray status table


```{r}
x <- x %>%
  dplyr::select(round, spray_status, Tests, positives, negatives, percent_positive)
prettify(x)
```  
 
Row
-------------------------------------


### Cumulative incidence figure
    
```{r}
#   Cumulative incidence
#   (Total # of RDT+ in the cohort since start / # of child-years at risk**)
#     ** number of children in the cohort x time they have been followed. If a child was LFU then he
#     contributed until that moment to child-time-at-risk
options(scipen = '999')
# Define a new case
etar$new_case <- etar$malaria & etar$visit
x <- etar %>%
  group_by(date) %>%
  summarise(children_at_risk = n(),
            new_cases = length(which(new_case))) %>%
  ungroup %>%
  mutate(cumulative_new_cases = cumsum(new_cases),
         cumulative_children_at_risk = cumsum(children_at_risk))  %>%
  mutate(cumulative_incidence = cumulative_new_cases / (cumulative_children_at_risk / 365.25))

# # total number of children years at risk
# denom <- dplyr::last(x$cumulative_children_at_risk) / 365.25
# 
# x <- x %>%
#   mutate(cumulative_incidence = cumulative_new_cases / denom)

ggplot(data = x %>% filter(date >= '2017-01-31'),
       aes(x = date,
           y = cumulative_incidence)) +
  geom_line() +
  theme_mopeia() +
  labs(x = '',
       y = 'Cumulative incidence per child-years at risk') 
``` 
    

### Cumulative incidence table

```{r}
prettify(x)
```

        
Row
-------------------------------------
    

    
### Cumulative incidence by spray status figure
    
```{r}
#     Cumulative incidence by spray status
#     As above but disaggregated by those in “spray” and “no-spray” areas.
#     Please add RR and 95%CI for the RR

x <- etar %>%
  group_by(date, spray_status) %>%
  summarise(children_at_risk = n(),
            new_cases = length(which(new_case))) %>%
  ungroup %>%
  group_by(spray_status) %>%
  mutate(cumulative_new_cases = cumsum(new_cases),
         cumulative_children_at_risk = cumsum(children_at_risk)) %>%
  ungroup %>%
  mutate(cumulative_incidence = cumulative_new_cases / (cumulative_children_at_risk / 365.25))

  # mutate(denom = dplyr::last(cumulative_children_at_risk) / 365.25) %>%
  # ungroup %>%
  # mutate(cumulative_incidence = cumulative_new_cases / denom)

ggplot(data = x %>% filter(date >= '2017-01-31'),
       aes(x = date,
           y = cumulative_incidence,
           color = spray_status)) +
  geom_line() +
  theme_mopeia() +
  labs(x = '',
       y = 'Cumulative incidence per child-years at risk') +
  scale_color_manual(name= '',
                     values = c('darkorange', 'blue'))
```

### Cumulative incidence by spray status table

```{r}
prettify(x)
```

    
Row
-------------------------------------
    

### Corrected incidence: removing 30 days from time-at-risk for every incident case figure

```{r}
library(RcppRoll)
#     Corrected incidence 1
#     Cumulative incidence BUT every time a child is positive we should remove 30 days of time-at-risk from
#     the denominator

x <- etar %>%
  mutate(val = as.numeric(new_case)) %>%
  # group_by(permid) %>%
  mutate(remove = RcppRoll::roll_sum(x = val, n = 30, fill = 0, na.rm = FALSE)) %>%
  ungroup %>%
  filter(remove == 0 | new_case) %>%
  group_by(date, spray_status) %>%
  summarise(children_at_risk = n(),
            new_cases = length(which(new_case))) %>%
  ungroup %>%
  group_by(spray_status) %>%
  mutate(cumulative_new_cases = cumsum(new_cases),
         cumulative_children_at_risk = cumsum(children_at_risk)) %>%
  mutate(cumulative_incidence = cumulative_new_cases / (cumulative_children_at_risk / 365.25))

ggplot(data = x %>% filter(date >= '2017-01-31'),
       aes(x = date,
           y = cumulative_incidence,
           color = spray_status)) +
  geom_line() +
  theme_mopeia() +
  labs(x = '',
       y = 'Cumulative incidence per child-years at risk') +
  scale_color_manual(name= '',
                     values = c('darkorange', 'blue'))

```
    
### Corrected incidence: removing 30 days from time-at-risk for every incident case table
    
    
```{r}
prettify(x)
```
    
Row
-------------------------------------

### Corrected incidence: removing 30 days for every incident case, and ignoring next visit figure

```{r}
x <- etar %>%
  mutate(val = as.numeric(new_case)) %>%
  # group_by(permid) %>%
  mutate(remove = RcppRoll::roll_sum(x = val, n = 30, fill = 0, na.rm = FALSE)) %>%
  ungroup %>%
  filter(remove == 0 | new_case)
# Identify next cases to be removed too
next_cases <- etar %>%
  filter(visit) %>%
  group_by(permid) %>%
  mutate(ignore = dplyr::lag(malaria, 1, default = FALSE)) %>%
  dplyr::select(permid, date, ignore)
  
x <- left_join(x, next_cases)
x <- x %>% filter(!ignore)
x <- x %>%
  group_by(date, spray_status) %>%
  summarise(children_at_risk = n(),
            new_cases = length(which(new_case))) %>%
  ungroup %>%
  group_by(spray_status) %>%
  mutate(cumulative_new_cases = cumsum(new_cases),
         cumulative_children_at_risk = cumsum(children_at_risk)) %>%
  mutate(cumulative_incidence = cumulative_new_cases / (cumulative_children_at_risk / 365.25))

ggplot(data = x %>% filter(date >= '2017-01-31'),
       aes(x = date,
           y = cumulative_incidence,
           color = spray_status)) +
  geom_line() +
  theme_mopeia() +
  labs(x = '',
       y = 'Cumulative incidence per child-years at risk') +
  scale_color_manual(name= '',
                     values = c('darkorange', 'blue'))
```

### Corrected incidence: removing 30 days for every incident case, and ignoring next visit table

```{r}
prettify(x)
```

Row
-------------------------------------
    

### Cumulative incidence by spray status and net usage (using variable `main_malaria_control_undernet_sleep`), no denominator removals figure

```{r}
#     As above but also disaggregated by sleep last night (question 9 of follow questionnaire)
nets <- acd %>% dplyr::select(perm_id, main_malaria_control_undernet_sleep) %>%
  rename(permid = perm_id) %>%
  rename(net = main_malaria_control_undernet_sleep)
# Decode
net_dict <- data_frame(net = as.character(1:5),
                       bednet = c('< 1 time per week',
                                  '1 time per week',
                                  '2-3 times per week',
                                  '4 or more times per week',
                                  'Respondent does not know'))
nets <- left_join(nets, net_dict)

# Join to etar

x <- etar %>%
  left_join(nets, by ='permid') %>%
  group_by(date, spray_status, bednet) %>%
  summarise(children_at_risk = n(),
            new_cases = length(which(new_case))) %>%
  ungroup %>%
  group_by(spray_status, bednet) %>%
  mutate(cumulative_new_cases = cumsum(new_cases),
         cumulative_children_at_risk = cumsum(children_at_risk)) %>%
  ungroup %>%
  mutate(cumulative_incidence = cumulative_new_cases / (cumulative_children_at_risk / 365.25)) #%>%
  # mutate(bednet = paste0('Bednet: ', bednet))

  # mutate(denom = dplyr::last(cumulative_children_at_risk) / 365.25) %>%
  # ungroup %>%
  # mutate(cumulative_incidence = cumulative_new_cases / denom)

ggplot(data = x %>% filter(date >= '2017-01-31'),
       aes(x = date,
           y = cumulative_incidence,
           color = bednet)) +
  geom_line() +
  theme_mopeia() +
  labs(x = '',
       y = 'Cumulative incidence per child-years at risk (log)') +
  scale_color_manual(name= 'Net',
                     values = rev(colorRampPalette(brewer.pal(n = 9, 'Spectral'))(length(unique(x$bednet))))) +
  facet_wrap(~spray_status) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(strip.text.x = element_text(size = 8)) +
  scale_y_log10()
```

### Cumulative incidence by spray status and net usage (using variable `main_malaria_control_undernet_sleep`), no denominator removals table

```{r}
prettify(x)
```


PCD
=====================================  
    
Row
-------------------------------------
    
### Monthly PCD malaria cases figure (0-4 years only)
    
```{r}
x <- pcd %>%
  dplyr::mutate(month = date_truncate(date, level = 'month')) %>%
  dplyr::group_by(month) %>%
  dplyr::summarise(cases = length(which(malaria)),
            valid_tests = length(which(!is.na(rdt))),
            visits = n()) %>%
  ungroup %>%
  dplyr::filter(month >= '2016-12-01')

ggplot(data = x,
       aes(x = month,
           y = cases)) +
  geom_bar(fill = 'darkorange',
           alpha = 0.6,
           stat = 'identity') +
  theme_mopeia() +
  labs(x = '',
       y = 'Positive RDTs')
```

### Monthly PCD malaria cases table (0-4 years only)

```{r}
prettify(x %>%
           mutate(month = format(month, '%B %Y')))
```

Row
-------------------------------------
   

### Monthly PCD malaria cases by health facility figure (0-4 years only)

```{r}
x <- pcd %>%
  dplyr::mutate(month = date_truncate(date, level = 'month')) %>%
  dplyr::group_by(month, health_facility) %>%
  dplyr::summarise(cases = length(which(malaria)),
            valid_tests = length(which(!is.na(rdt))),
            visits = n()) %>%
  ungroup %>%
  dplyr::filter(month >= '2016-12-01')

ggplot(data = x,
       aes(x = month,
           y = cases)) +
  geom_bar(fill = 'darkorange',
           alpha = 0.6,
           stat = 'identity') +
  theme_mopeia() +
  labs(x = '',
       y = 'Positive RDTs') +
  facet_wrap(~health_facility)
```

### Monthly PCD malaria cases by health facility table (0-4 years only)


```{r}
prettify(x %>%
           mutate(month = format(month, '%B %Y')))
```

Row
-------------------------------------


### Monthly PCD malaria cases by spray status figure (0-4 years only)

```{r}
x <- pcd %>%
  dplyr::mutate(month = date_truncate(date, level = 'month')) %>%
  dplyr::group_by(month, spray_status) %>%
  dplyr::summarise(cases = length(which(malaria)),
            valid_tests = length(which(!is.na(rdt))),
            visits = n()) %>%
  ungroup %>%
  dplyr::filter(month >= '2016-12-01') %>%
  mutate(spray_status = ifelse(is.na(spray_status),
                               'Unknown',
                               spray_status))

ggplot(data = x,
       aes(x = month,
           y = cases,
           color = spray_status)) +
  geom_line(alpha = 0.6,
           stat = 'identity') +
  geom_point() +
  theme_mopeia() +
  labs(x = '',
       y = 'Positive RDTs') +
  scale_x_date(breaks = sort(unique(x$month))) +
  scale_color_manual(name= '',
                     values = c('darkorange', 'blue', 'black')) +
  theme(axis.text.x = element_text(angle = 90))
```

### Monthly PCD malaria cases by spray status table (0-4 years only)


```{r}
prettify(x %>%
           mutate(month = format(month, '%B %Y')))
```
   
Row
-------------------------------------



### Cumulative monthly PCD malaria cases by spray status figure (0-4 years only)

```{r}
x <- pcd %>%
  dplyr::mutate(month = date_truncate(date, level = 'month')) %>%
  dplyr::group_by(month, spray_status) %>%
  dplyr::summarise(cases = length(which(malaria)),
            valid_tests = length(which(!is.na(rdt))),
            visits = n()) %>%
  ungroup %>%
  dplyr::filter(month >= '2016-12-01') %>%
  mutate(spray_status = ifelse(is.na(spray_status),
                               'Unknown',
                               spray_status)) %>%
  arrange(month) %>%
  group_by(spray_status) %>%
  mutate(cumulative_cases = cumsum(cases)) %>%
  ungroup

ggplot(data = x,
       aes(x = month,
           y = cumulative_cases,
           color = spray_status)) +
  geom_line(alpha = 0.6,
           stat = 'identity') +
  geom_point() +
  theme_mopeia() +
  labs(x = '',
       y = 'Positive RDTs') +
  scale_x_date(breaks = sort(unique(x$month))) +
  scale_color_manual(name= '',
                     values = c('darkorange', 'blue', 'black')) +
  theme(axis.text.x = element_text(angle = 90))
```

### Cumulative monthly PCD malaria cases by spray status table (0-4 years only)


```{r}
prettify(x %>%
           mutate(month = format(month, '%B %Y')) %>%
           dplyr::select(month, spray_status, cumulative_cases))
```



Analysis
=====================================  

Row
-------------------------------------
    
### Month-specific relative risk from ACD (spray relative to no spray) figure

```{r}
library("epitools")

x <- etar %>%    
  mutate(month = date_truncate(date, 'month')) %>%
  group_by(month, spray_status) %>%
  summarise(children_at_risk = n(),
            new_cases = length(which(new_case & visit))) %>%
  ungroup %>%
  dplyr::select(month, spray_status, new_cases, children_at_risk) %>%
  group_by(month) %>%
  filter(n() == 2) %>%
  summarise(cases_spray = new_cases[spray_status == 'Spray'],
            cases_no_spray = new_cases[spray_status == 'No spray'],
            at_risk_spray = children_at_risk[spray_status == 'Spray'],
            at_risk_no_spray = children_at_risk[spray_status == 'No spray'])
# Calculate relative risk and confidence intervals
x$rr <- x$lower <- x$upper <- x$p <- NA
for (i in 1:nrow(x)){
  tab <- matrix(as.numeric(x[i,2:5]), nrow = 2)
  out <- epitab(tab,method="riskratio")$tab
  x$rr[i] <- out[,colnames(out) == 'riskratio'][2]
  x$lower[i] <- out[,colnames(out) == 'lower'][2]
  x$upper[i] <- out[,colnames(out) == 'upper'][2]
  x$p[i] <- out[,colnames(out) == 'p.value'][2]
}

ggplot(data = x,
       aes(x = month,
           y = rr)) +
  geom_point() +
  geom_linerange(aes(ymax = upper,
                     ymin = lower)) +
  theme_mopeia() +
  labs(x = '',
       y = 'Relative risk') +
  scale_x_date(breaks = sort(unique(x$month)),
               date_labels = '%b') +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_hline(yintercept = 1,
             color = 'red')

```

### Month-specific relative risk from ACD (spray relative to no spray) table

```{r}
prettify(x)
```
   
    
Row
-------------------------------------
    
### Cumulative relative risk from ACD (spray relative to no spray) figure


```{r}
library("epitools")

x <- etar %>%
  group_by(date, spray_status) %>%
  summarise(children_at_risk = n(),
            new_cases = length(which(new_case & visit))) %>%
  ungroup %>%
  group_by(spray_status) %>%
  mutate(cumulative_new_cases = cumsum(new_cases),
         cumulative_children_at_risk = cumsum(children_at_risk))  %>%
  mutate(cumulative_incidence = cumulative_new_cases / (cumulative_children_at_risk / 365.25)) %>%
  ungroup %>%
  mutate(month = date_truncate(date, 'month')) %>%
  group_by(month) %>%
  filter(date == max(date)) %>%
  ungroup %>%
  dplyr::select(date, spray_status, cumulative_new_cases, 
                cumulative_children_at_risk) %>%
  group_by(date) %>%
  filter(n() == 2) %>%
  summarise(cases_spray = cumulative_new_cases[spray_status == 'Spray'],
            cases_no_spray = cumulative_new_cases[spray_status == 'No spray'],
            at_risk_spray = cumulative_children_at_risk[spray_status == 'Spray'],
            at_risk_no_spray = cumulative_children_at_risk[spray_status == 'No spray'])
# Calculate relative risk and confidence intervals
x$rr <- x$lower <- x$upper <- x$p <- NA
for (i in 1:nrow(x)){
  tab <- matrix(as.numeric(x[i,2:5]), nrow = 2)
  out <- epitab(tab,method="riskratio")$tab
  x$rr[i] <- out[,colnames(out) == 'riskratio'][2]
  x$lower[i] <- out[,colnames(out) == 'lower'][2]
  x$upper[i] <- out[,colnames(out) == 'upper'][2]
  x$p[i] <- out[,colnames(out) == 'p.value'][2]
}

ggplot(data = x,
       aes(x = date,
           y = rr)) +
  geom_point() +
  geom_linerange(aes(ymax = upper,
                     ymin = lower)) +
  theme_mopeia() +
  labs(x = '',
       y = 'Relative risk') +
  scale_x_date(breaks = sort(unique(x$date)),
               date_labels = '%b') +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_hline(yintercept = 1,
             color = 'red')
```

    
### Cumulative relative risk from ACD (spray relative to no spray) table

```{r}
prettify(x)
```



Row
-------------------------------------
    
### Month-specific odds ratios from ACD (spray relative to no spray) figure

```{r}
library("epitools")

x <- etar %>%    
  mutate(month = date_truncate(date, 'month')) %>%
  group_by(month, spray_status) %>%
  summarise(children_at_risk = n(),
            new_cases = length(which(new_case & visit))) %>%
  ungroup %>%
  dplyr::select(month, spray_status, new_cases, children_at_risk) %>%
  group_by(month) %>%
  filter(n() == 2) %>%
  summarise(cases_spray = new_cases[spray_status == 'Spray'],
            cases_no_spray = new_cases[spray_status == 'No spray'],
            at_risk_spray = children_at_risk[spray_status == 'Spray'],
            at_risk_no_spray = children_at_risk[spray_status == 'No spray'])
# Calculate odds ratios and confidence intervals
x$rr <- x$lower <- x$upper <- x$p <- NA
for (i in 1:nrow(x)){
  tab <- matrix(as.numeric(x[i,2:5]), nrow = 2)
  out <- epitab(tab,method="oddsratio")$tab
  x$rr[i] <- out[,colnames(out) == 'oddsratio'][2]
  x$lower[i] <- out[,colnames(out) == 'lower'][2]
  x$upper[i] <- out[,colnames(out) == 'upper'][2]
  x$p[i] <- out[,colnames(out) == 'p.value'][2]
}

ggplot(data = x,
       aes(x = month,
           y = rr)) +
  geom_point() +
  geom_linerange(aes(ymax = upper,
                     ymin = lower)) +
  theme_mopeia() +
  labs(x = '',
       y = 'odds ratios') +
  scale_x_date(breaks = sort(unique(x$month)),
               date_labels = '%b') +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_hline(yintercept = 1,
             color = 'red')

```

### Month-specific odds ratios from ACD (spray relative to no spray) table

```{r}
prettify(x)
```
   
    
Row
-------------------------------------
    
### Cumulative odds ratios from ACD (spray relative to no spray) figure


```{r}
library("epitools")

x <- etar %>%
  group_by(date, spray_status) %>%
  summarise(children_at_risk = n(),
            new_cases = length(which(new_case & visit))) %>%
  ungroup %>%
  group_by(spray_status) %>%
  mutate(cumulative_new_cases = cumsum(new_cases),
         cumulative_children_at_risk = cumsum(children_at_risk))  %>%
  mutate(cumulative_incidence = cumulative_new_cases / (cumulative_children_at_risk / 365.25)) %>%
  ungroup %>%
  mutate(month = date_truncate(date, 'month')) %>%
  group_by(month) %>%
  filter(date == max(date)) %>%
  ungroup %>%
  dplyr::select(date, spray_status, cumulative_new_cases, 
                cumulative_children_at_risk) %>%
  group_by(date) %>%
  filter(n() == 2) %>%
  summarise(cases_spray = cumulative_new_cases[spray_status == 'Spray'],
            cases_no_spray = cumulative_new_cases[spray_status == 'No spray'],
            at_risk_spray = cumulative_children_at_risk[spray_status == 'Spray'],
            at_risk_no_spray = cumulative_children_at_risk[spray_status == 'No spray'])
# Calculate odds ratios and confidence intervals
x$rr <- x$lower <- x$upper <- x$p <- NA
for (i in 1:nrow(x)){
  tab <- matrix(as.numeric(x[i,2:5]), nrow = 2)
  out <- epitab(tab,method="oddsratio")$tab
  x$rr[i] <- out[,colnames(out) == 'oddsratio'][2]
  x$lower[i] <- out[,colnames(out) == 'lower'][2]
  x$upper[i] <- out[,colnames(out) == 'upper'][2]
  x$p[i] <- out[,colnames(out) == 'p.value'][2]
}

ggplot(data = x,
       aes(x = date,
           y = rr)) +
  geom_point() +
  geom_linerange(aes(ymax = upper,
                     ymin = lower)) +
  theme_mopeia() +
  labs(x = '',
       y = 'odds ratios') +
  scale_x_date(breaks = sort(unique(x$date)),
               date_labels = '%b') +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_hline(yintercept = 1,
             color = 'red')
```

    
### Cumulative odds ratios from ACD (spray relative to no spray) table

```{r}
prettify(x)
```



Row
-------------------------------------
    
### Month-specific odds ratios from PCD (spray relative to no spray) 

```{r}
library("epitools")

# Get denominators
by_ss <- master_table %>%
  group_by(spray_status = ifelse(status, 'Spray', 'No spray')) %>%
  summarise(children = n())

x <- pcd %>%    
  # mutate(month = date_truncate(date, 'month')) %>%
  # group_by(month, spray_status) %>%
  group_by(date, spray_status) %>%
  summarise(new_cases = length(which(malaria))) %>%
  ungroup %>%
  filter(!is.na(spray_status))
left <- expand.grid(date = seq(as.Date('2016-12-01'), 
                              max(x$date),
                              by = 1),
                    spray_status = sort(unique(x$spray_status)))
x <- left_join(left,
               x)
x$new_cases[is.na(x$new_cases)] <- 0
x <- x %>% filter(!is.na(spray_status))  

# Join denominators
x <- x %>%
  left_join(by_ss)

x <- x %>%
  dplyr::arrange(date) %>%
  dplyr::group_by(spray_status) %>%
  dplyr::mutate(cumulative_new_cases = cumsum(new_cases),
            cumulative_children_at_risk = cumsum(children)) %>%
  ungroup %>%
  mutate(month = date_truncate(date, 'month')) %>%
  # group_by(month) %>%
  # filter(date == max(date)) %>%
  # ungroup %>%
  rename(children_at_risk = children) %>%
  group_by(month) %>%
  # filter(n() == 2) %>%
  summarise(cases_spray = sum(new_cases[spray_status == 'Spray']),
            cases_no_spray = sum(new_cases[spray_status == 'No spray']),
            at_risk_spray = sum(children_at_risk[spray_status == 'Spray']),
            at_risk_no_spray = sum(children_at_risk[spray_status == 'No spray']))
# Calculate odds ratios and confidence intervals
x$rr <- x$lower <- x$upper <- x$p <- NA
for (i in 1:nrow(x)){
  tab <- matrix(as.numeric(x[i,2:5]), nrow = 2)
  out <- epitab(tab,method="oddsratio")$tab
  x$rr[i] <- out[,colnames(out) == 'oddsratio'][2]
  x$lower[i] <- out[,colnames(out) == 'lower'][2]
  x$upper[i] <- out[,colnames(out) == 'upper'][2]
  x$p[i] <- out[,colnames(out) == 'p.value'][2]
}

ggplot(data = x,
       aes(x = month,
           y = rr)) +
  geom_point() +
  geom_linerange(aes(ymax = upper,
                     ymin = lower)) +
  theme_mopeia() +
  labs(x = '',
       y = 'odds ratios') +
  scale_x_date(breaks = sort(unique(x$month)),
               date_labels = '%b') +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_hline(yintercept = 1,
             color = 'red')
```

    
### Month-specific odds ratios from PCD (spray relative to no spray) table

```{r}
prettify(x)
```

Row
-------------------------------------
  
      
### Cumulative odds ratios from PCD (spray relative to no spray) 

```{r}
library("epitools")

# Get denominators
by_ss <- master_table %>%
  group_by(spray_status = ifelse(status, 'Spray', 'No spray')) %>%
  summarise(children = n())

x <- pcd %>%    
  # mutate(month = date_truncate(date, 'month')) %>%
  # group_by(month, spray_status) %>%
  group_by(date, spray_status) %>%
  summarise(new_cases = length(which(malaria))) %>%
  ungroup %>%
  filter(!is.na(spray_status))
left <- expand.grid(date = seq(as.Date('2016-12-01'), 
                              max(x$date),
                              by = 1),
                    spray_status = sort(unique(x$spray_status)))
x <- left_join(left,
               x)
x$new_cases[is.na(x$new_cases)] <- 0
x <- x %>% filter(!is.na(spray_status))  

# Join denominators
x <- x %>%
  left_join(by_ss)

x <- x %>%
  dplyr::arrange(date) %>%
  dplyr::group_by(spray_status) %>%
  dplyr::mutate(cumulative_new_cases = cumsum(new_cases),
            cumulative_children_at_risk = cumsum(children)) %>%
  ungroup %>%
  mutate(month = date_truncate(date, 'month')) %>%
  group_by(month) %>%
  filter(date == max(date)) %>%
  ungroup %>%
  rename(children_at_risk = children) %>%
  group_by(month) %>%
  # filter(n() == 2) %>%
  summarise(cases_spray = cumulative_new_cases[spray_status == 'Spray'],
            cases_no_spray = cumulative_new_cases[spray_status == 'No spray'],
            at_risk_spray = cumulative_children_at_risk[spray_status == 'Spray'],
            at_risk_no_spray = cumulative_children_at_risk[spray_status == 'No spray'])
# Calculate odds ratios and confidence intervals
x$rr <- x$lower <- x$upper <- x$p <- NA
for (i in 1:nrow(x)){
  tab <- matrix(as.numeric(x[i,2:5]), nrow = 2)
  out <- epitab(tab,method="oddsratio")$tab
  x$rr[i] <- out[,colnames(out) == 'oddsratio'][2]
  x$lower[i] <- out[,colnames(out) == 'lower'][2]
  x$upper[i] <- out[,colnames(out) == 'upper'][2]
  x$p[i] <- out[,colnames(out) == 'p.value'][2]
}

ggplot(data = x,
       aes(x = month,
           y = rr)) +
  geom_point() +
  geom_linerange(aes(ymax = upper,
                     ymin = lower)) +
  theme_mopeia() +
  labs(x = '',
       y = 'odds ratios') +
  scale_x_date(breaks = sort(unique(x$month)),
               date_labels = '%b') +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_hline(yintercept = 1,
             color = 'red')
```

    
### Cumulative odds ratios from PCD (spray relative to no spray) table

```{r}
prettify(x)
```

  


Details
=====================================  

Row
-------------------------------------
  

### Data management

```{r}
valueBox('Eldo Elobolobo', icon = "fa-database",
         href="mailto:eldo.elobolobo@manhica.net?Subject=Mopeia%20dashboard")
```

### Visualization

```{r}
valueBox('Joe Brew', icon = "fa-eye",
         href="mailto:joe@databrew.cc?Subject=Mopeia%20dashboard")
```

### Implementation and analysis

```{r}
valueBox('Carlos Chaccour', icon = "fa-stethoscope",
         href="mailto:carlos.chaccour@isglobal.org?Subject=Mopeia%20dashboard")
```


Row
-------------------------------------
  

### Coordinating institution

```{r}
valueBox('Centro de Investigação em Saude de Manhiça', icon = "fa-users",
         href="http://www.manhica.org/wp/en/home-en/", color = 'danger')
```

Row
-------------------------------------
  

### Collaborating institution

```{r}
valueBox('Barcelona Institute for Global Health', icon = "fa-group",
         href="http://www.isglobal.org", color = 'danger')
```


Row
-------------------------------------
  
### Supporting organization

```{r}
valueBox('DataBrew', icon = "fa-coffee",
         href="http://www.databrew.cc", color = 'danger')
```
